---
title: "Basic usage"
author: "Mauricio Vargas S."
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to use this package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  message = FALSE,
  comment = "#>"
)

datatable <- function(x) {
  DT::datatable(x,
    extensions = "FixedColumns",
    options = list(
      pageLength = 5,
      dom = 'Bfrtip',
      scrollX = TRUE,
      fixedColumns = list(leftColumns = 2, rightColumns = 1)
    )
)}
```

# Introduction

This vignette explains the functions within this package. The idea is to show how this package simplifies obtaining data from (api.tradestatistics.io)[https://api.tradestatistics.io].

To improve the presentation of the tables I shall use `DT` and the `datatable()` function besides `tradestatistics`.
```{r pkgs}
library(tradestatistics)
library(DT)
```

# Package data

## Available tables

Provided that this package obtains data from an API, it is useful to know which tables can be accessed:

```{r tables, eval = T}
datatable(ots_tables)
```

You might notice the tables have a pattern. The letters indicate the presence of columns that account for the level of detail in the data:

* `y`: *y*ear column.
* `r`: *r*eporter column
* `p`: *p*artner column
* `c`: *c*ommodity column

The most aggregated table is `yr` which basically says how many dollars each country exports and imports for a given year.

The less aggregated table is `yrpc` which says how many dollars of each of the 1,242 commodities from the Harmonized System each country exports to other countries and imports from other countries.

For the complete detail you can check [tradestatistics.io](https://tradestatistics.io).

## Country codes

The Package Functions section explains that you don't need to memorize all ISO codes. The functions within this package are designed to match strings (i.e. "United States" or "America") to valid ISO codes (i.e. "USA").

Just as a reference, the table with all valid ISO codes can be accessed by running this:

```{r countries, eval = T}
datatable(ots_countries)
```

## Commodity codes

The Package Functions section explains that you don't need to memorize all HS codes. The functions within this package are designed to match strings (i.e. "apple") to valid HS codes (i.e. "0808").

```{r commodities, eval = T}
datatable(ots_commodities)
```

## Commodity shortnames

```{r commodities2, eval = T}
datatable(ots_commodities_shortnames)
```

## Commodity communities
```{r commodities3, eval = T}
datatable(ots_communities)
```
## Inflation data

This table is provided to be used with `ots_inflation_adjustment()`.

```{r inflation, eval = T}
datatable(ots_inflation)
```

# Package functions

## Country code

The end user can use this function to find an ISO code by providing a country name. This works by implementing partial search.

Basic examples:
```{r country_code}
# Single match with no replacement
datatable(ots_country_code("Chile"))

# Single match with replacement
datatable(ots_country_code("America"))

# Double match with no replacement
datatable(ots_country_code("Germany"))
```

The function `ots_country_code()` is used by `ots_create_tidy_data()` in a way that you can pass parameters like `ots_create_tidy_data(... reporters = "Chile" ...)` and it will automatically replace your input for a valid ISO in case there is a match. This will be covered in detail in the Trade Data section.

## Commodity code

The end user can find a code or a set of codes by looking for keywords for commodities or groups. The function `ots_commodity_code()` allows to search from the official commodities and groups in the Harmonized system:
```{r commodity_code2}
datatable(ots_commodity_code(commodity = " ShEEp ", group = " mEaT "))
```

## Community code

It's also possible to match commodities to communities (i.e. larger aggregation than community groups) created by the Center for International Development at Harvard:
```{r commodity_code}
datatable(ots_commodity_community(" ANIMAL "))
```

While there are 99 official community groups, the total number of commodity communities is 22. This is specially useful to visualize data but it not an official United Nations aggregation.

## Trade data

This function downloads data for a single year and needs (at least) some filter parameters according to the query type.

Here we cover aggregated tables to describe the usage.

### Bilateral trade at commodity level (Year - Reporter - Partner - Commodity Code)

If we want Chile-Argentina bilateral trade at community level in 1962:
```{r yrpc1, eval = T}
yrpc <- ots_create_tidy_data(
  years = 1990,
  reporters = "chl",
  partners = "arg",
  table = "yrpc"
)

datatable(yrpc)
```

We can pass two years or more, several reporters/partners, and filter by commodities with exact codes or code matching based on keywords:
```{r yrpc2, eval = T}
# Note that here I'm passing Peru and not per which is the ISO code for Peru
# The same applies to Brazil
yrpc2 <- ots_create_tidy_data(
  years = c(1962,1963),
  reporters = c("chl", "Peru", "bol"),
  partners = c("arg", "Brazil"),
  commodities = c("01", "food"),
  table = "yrpc"
)
datatable(yrpc2)
```

The `yrpc` table returns some fields that deserve an explanation which can be seen at [tradestatistics.io](https://tradestatistics.io). This example is interesting because "01" return a set of commodities (all commodities starting with 01, which is the commodity group "Animals; live"), but "food" return all commodities with a matching description ("1601", "1806", "1904", etc.). In addition, not all the requested commodities are exported from each reporter to each partner, therefore a warning is returned.

### Bilateral trade at aggregated level (Year - Reporter - Partner)

If we want Chile-Argentina bilateral trade at aggregated level in 1962 and 1963:

```{r yrp3, eval = T}
yrp <- ots_create_tidy_data(
  years = 1962:1963,
  reporters = c("chl", "per"),
  partners = "arg",
  table = "yrp"
)

datatable(yrp)
```

This table accepts different years, reporters and partners just like `yrpc`.

### Reporter trade at commodity level (Year - Reporter - Commodity Code) 

If we want Chilean trade at commodity level in 1962 with respect to commodity "0101" which means "Horses, asses, mules and hinnies; live":
```{r yrc2, eval = T}
yrc <- ots_create_tidy_data(
  years = 1962,
  reporters = "chl",
  commodities = "0101",
  table = "yrc"
)

datatable(yrc)
```

This table accepts different years, reporters and commodity codes just like `yrpc`.

All the variables from this table are documented at [tradestatistics.io](https://tradestatistics.io).

### Reporter trade at aggregated level (Year - Reporter)

If we want the aggregated trade of Chile, Argentina and Peru in 1962 and 1963:
```{r yr2, eval = T}
yr <- ots_create_tidy_data(
  years = 1962:1963,
  reporters = c("chl", "arg", "per"),
  table = "yr"
)

datatable(yr)
```

This table accepts different years and reporters just like `yrpc`.

All the variables from this table are documented at [tradestatistics.io](https://tradestatistics.io).

### Commodity trade at aggregated level (Year - Commodity Code)

If we want all commodities traded in 1962:
```{r yc1, eval = T}
yc <- ots_create_tidy_data(
  years = 1962,
  table = "yc"
)

datatable(yc)
```

If we want the traded values of the commodity "0808" which means "Horses, asses, mules and hinnies; live: in 1962:" in 1962:
```{r yc2, eval = T}
yc2 <- ots_create_tidy_data(
  years = 1962,
  commodities = "0808",
  table = "yc"
)

datatable(yc2)
```

This table accepts different years just like `yrpc`.

## Inflation adjustment

Taking the `yr` table from above, we can use `ots_inflation_adjustment()` to convert dollars from 1962 to dollars of 1970:

```{r}
inflation <- ots_inflation_adjustment(yr, reference_year = 1970)
datatable(inflation)
```
