---
title: "How to use this package"
author: "Mauricio Vargas S."
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to use this package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  message = FALSE,
  comment = "#>"
)
```

# Package data

## Available tables

Provided that this package obtains data from an API, it is useful to know which tables can be accessed:

```{r tables, eval = T}
library(tradestatistics)

ots_tables
```

You might notice the tables have a pattern. The letters indicate the presence of columns that account for the level of detail in the data:

* `y`: *y*ear column.
* `r`: *r*eporter column
* `p`: *p*artner column
* `c`: product *c*ode column
* `g`: product *g*roup column
* `o`: product c*o*mmunity column

The most aggregated table is `yr` which basically says how many dollars each country exports and imports for a given year.

The less aggregated table is `yrpc` which says how many dollars of each of the 1,222 products from the Harmonized System each country exports to other countries and imports from other countries.

`g` and `o` are aggregations over `c`, therefore `yrpc-ga` (`yrpc` with group aggregation) and `yrpc-ga` (`yrpc` with product community aggregation) are summaries of `yrpc`.

For the complete detail you can check [docs.tradestatistics.io](docs.tradestatistics.io).

## Country codes

The Package Functions section explains that you don't need to memorize all ISO codes. The functions within this package are designed to match strings (i.e. "United States" or "America") to valid ISO codes (i.e. "USA").

Just as a reference, the table with all valid ISO codes can be accessed by running this:

```{r countries, eval = T}
ots_countries
```

## Product codes and communities

The Package Functions section explains that you don't need to memorize all HS codes. The functions within this package are designed to match strings (i.e. "apple") to valid HS codes (i.e. "0808").

```{r products, eval = T}
ots_products

ots_product_shortnames

ots_communities
```

## Inflation data

This table is provided to be used with `ots_inflation_adjustment()`.

```{r inflation, eval = T}
ots_inflation
```

# Package functions

## Country code

The end user can use this function to find an ISO code by providing a country name. This works by implementing partial search.

Basic examples:
```{r country_code}
# Single match with no replacement
ots_country_code("Chile")

# Single match with replacement
ots_country_code("America")

# Double match with no replacement
ots_country_code("Germany")
```

The function `ots_country_code()` is used by `ots_create_tidy_data()` in a way that you can pass parameters like `ots_create_tidy_data(... reporters = "Chile" ...)` and it will automatically replace your input for a valid ISO in case there is a match. This will be covered in detail in the Trade Data section.

## Product code

The end user can use this function to find a product code by providing a product name. This works by implementing partial string matching:

```{r product_code}
ots_product_code("wine")
```

This function searches both products and groups:
```{r product_code2}
ots_product_code(productname = "Milk", productgroup = "Dairy")
```

## Community code

The end user can use this function to find a community code by providing a community name. This works by implementing partial string matching:

```{r community_code}
ots_product_community("fabric")
```

## Trade data

This function downloads data for a single year and needs (at least) some filter parameters according to the query type.

Here we cover aggregated tables to describe the usage.

### Bilateral trade at community level (Year - Reporter - Partner - Community Code)

If we want Chile-Argentina bilateral trade at community level in 1962:
```{r yrpc-ga1, eval = T}
ots_create_tidy_data(
  years = 1962,
  reporters = "chl",
  partners = "arg",
  table = "yrpc-ca"
)
```

We can pass two years or more, several reporters/partners, and filter by community with exact codes or code matching based on keywords:
```{r yrpc-ga2, eval = T}
# Note that here I'm passing Peru and not per which is the ISO code for Peru
# The same applies to Brazil
ots_create_tidy_data(
  years = c(1962,1963),
  reporters = c("chl", "Peru", "bol"),
  partners = c("arg", "Brazil"),
  communities = c("01", "food"),
  table = "yrpc-ca"
)
```

`yrpc-ca` table returns some fields that deserve an explanation which can be seen at [docs.tradestatistics.io](docs.tradestatistics.io).

`yrpc-ga` and `yrpc` tables are less aggregated version of this, and the usage is almost the same.

### Bilateral trade at aggregated level (Year - Reporter - Partner)

If we want Chile-Argentina bilateral trade at aggregated level in 1962 and 1963:

```{r yrp3, eval = T}
ots_create_tidy_data(
  years = 1962:1963,
  reporters = c("chl", "per"),
  partners = "arg",
  table = "yrp"
)
```

This table accepts different years, reporters and partners just like `yrpc-ca`.

### Reporter trade at product level (Year - Reporter - Product Code) 

If we want Chilean trade at product level in 1962 with respect to product `0101` which means "Horses, asses, mules and hinnies; live":
```{r yrc2, eval = T}
ots_create_tidy_data(
  years = 1962,
  reporters = "chl",
  products = "0101",
  table = "yrc"
)
```

This table accepts different years, reporters and product codes just like `yrpc-ca`.

All the variables from this table are documented at [docs.tradestatistics.io](docs.tradestatistics.io).

### Reporter trade at aggregated level (Year - Reporter)

If we want the aggregated trade of Chile, Argentina and Peru in 1962 and 1963:
```{r yr2, eval = T}
yr <- ots_create_tidy_data(
  years = 1962:1963,
  reporters = c("chl", "arg", "per"),
  table = "yr"
)

yr
```

This table accepts different years and reporters just like `yrpc`.

All the variables from this table are documented at [docs.tradestatistics.io](docs.tradestatistics.io).

### Product trade at aggregated level (Year - Product Code)

If we want all products traded in 1962:
```{r yc1, eval = T}
ots_create_tidy_data(
  years = 1962,
  table = "yc"
)
```

If we want the traded values of the product `0101` which means "Horses, asses, mules and hinnies; live: in 1962:
```{r yc2, eval = T}
ots_create_tidy_data(
  years = 1962,
  products = "0101",
  table = "yc"
)
```

This table accepts different years just like `yrpc`.

## Inflation adjustment

Taking the `yr` table from above, we can use `ots_inflation_adjustment()` to convert dollars from 1962 to dollars of 1970:

```{r}
ots_inflation_adjustment(yr, reference_year = 1970)
```
