---
title: "How to use this package"
author: "Mauricio Vargas S."
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to use this package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

# Package data

## Available tables

Provided that this package obtains data from an API, it is useful to know which tables can be accessed:

```{r tables, eval = T}
library(tradestatistics)

ots_attributes_tables
```

## Country codes

The Package Functions section explains that you don't need to memorize all ISO codes. The functions within this package are designed to match strings (i.e. "United States" or "America") to valid ISO codes (i.e. "USA").

Just as a reference, the table with all valid ISO codes can be accessed by running this:

```{r countries, eval = T}
ots_attributes_countries
```

## Product codes

The Package Functions section explains that you don't need to memorize all HS codes. The functions within this package are designed to match strings (i.e. "apple") to valid HS codes (i.e. "0808").

```{r products, eval = T}
ots_attributes_countries
```

# Package functions

## Country code

The end user can use this function to find an ISO code by providing a country name. This works by implementing partial search.

Basic examples:
```{r country_code}
# Single match with no replacement
ots_country_code("Chile")

# Single match with replacement
ots_country_code("America")

# Double match with no replacement
ots_country_code("Germany")
```

The function `ots_country_code()` is used by `ots_create_tidy_data()` in a way that you can pass parameters like `ots_create_tidy_data(... reporters = "Chile" ...)` and it will automatically replace your input for a valid ISO in case there is a match. This will be covered in detail in the Trade Data section.

## Product code

The end user can use this function to find a product code by providing a product name. This works by implementing partial string matching:

```{r product_code}
ots_product_code("wine")
```

## Trade data

This function downloads data for a single year and needs (at least) some filter parameters according to the query type.
 
### Bilateral trade at product level (Year - Reporter - Partner - Product Code)

If we want Chile-Argentina bilateral trade at product level in 1965:
```{r yrpc1, eval = T}
ots_create_tidy_data(years = 1965, reporters = "chl", partners = "arg")

# the same can be obtained specifying yrpc which is the default table
# ots_create_tidy_data(years = 1965, reporters = "chl", partners = "arg", table = "yrpc")
```

We can pass more than one year and or reporter/partner:
```{r yrpc2, eval = T}
# Note that here I'm passing Peru and not per which is the ISO code for Peru
# The same applies to Brazil
ots_create_tidy_data(years = c(1962,1965), reporters = c("chl", "Peru", "bol"), partners = c("arg", "Brazil"))
```

If we want filtering by product, there are different options:
```{r yrpc3, eval = T}
# Pass a specific HS code
ots_create_tidy_data(years = c(1962,1965), reporters = c("chl", "Peru", "bol"), partners = c("arg", "bra"), products = "0101")

# Pass a string that will return all matching descriptions and multiple HS codes
ots_create_tidy_data(years = c(1962,1965), reporters = c("chl", "Peru", "bol"), partners = c("arg", "bra"), products = c("0101", "apple"))
```

`yrpc` table returns some fields that deserve an explanation:

* `product_code`: HS07 product codes (e.g. according to the \code{products} table within this package, 0101 stands for "Horses, etc.")
* `product_code_length`: How many digits does `product_code` contain, this can be useful to filter by depth when using HS codes (HS 6 digits is a more detailed version of HS 4 digits, and therefore you don't have to sum both or you'll be counting exports/imports twice)
* `group_code`: International categorization of group products defined after product ID
* `group_name`: English name corresponding to `group_id`
* `export_value_usd`: Exports measured in nominal United States Dollars (USD)
* `import_value_usd`: Imports measured in nominal United States Dollars (USD)
* `export_value_usd_percentage_change_1_year`: Nominal increase/decrease in exports measured as percentage with respect to last year
* `export_value_usd_percentage_change_5_years`: Nominal increase/decrease in exports measured as percentage with respect to five years ago
* `export_value_usd_change_1_year`: Nominal increase/decrease in exports measured in USD with respect to last year
* `export_value_usd_change_5_years`: Nominal increase/decrease in exports measured in USD with respect to five years ago

### Bilateral trade at aggregated level (Year - Reporter - Partner)

If we want Chile-Argentina bilateral trade at aggregated level in 1965:
```{r yrp1, eval = T}
ots_create_tidy_data(years = 1965, reporters = "chl", partners = "arg", table = "yrp")
```

Another option is to pass more than one year, reporter and/or partner:
```{r yrp3, eval = T}
ots_create_tidy_data(years = 1962:1965, reporters = c("chl", "per"), partners = "arg", table = "yrp")
```

This table accepts different years, reporters and partners just like `yrpc`.

### Reporter trade at product level (Year - Reporter - Product Code) 

If we want Chilean trade at product level in 1965:
```{r yrc1, eval = T}
ots_create_tidy_data(years = 1965, reporters = "chl", table = "yrc")
```

If we want Chilean trade at product level in 1965 with respect to product "0101" (Horses, asses, mules and hinnies; live):
```{r yrc2, eval = T}
ots_create_tidy_data(years = 1965, reporters = "chl", products = "0101", table = "yrc")
```

This table accepts different years, reporters and product codes just like `yrpc`.

Here the `export_rca*` and `import_rca*` fields contain the Revealed Comparative Advantage (RCA) of an exported product with respect to all the products with the same number of digits. The definition of RCA is detailed on [Open Trade Statistics Documentation](https://docs.tradestatistics.io/).

### Reporter trade at aggregated level (Year - Reporter)

If we want Chilean trade at aggregated level in 1965:
```{r yr, eval = T}
ots_create_tidy_data(years = 1965, reporters = "chl", table = "yr")
```

Another option is to pass more than one year and/or reporter:
```{r yr2, eval = T}
ots_create_tidy_data(years = 1962:1965, reporters = c("chl", "arg", "per"), table = "yr")
```

This table accepts different years and reporters just like `yrpc`.

Here some fields deserve an explanation:

* `eci_4_digits_product_code`: Economic Complexity Index (ECI) which is detailed on [Open Trade Statistics Documentation](https://docs.tradestatistics.io/). This index is built by using just four digits product codes.
* `eci_rank_4_digits_product_code`: The rank of a country given its ECI (e.g. the highest ECI obtains the #1)
* `eci_rank_4_digits_product_code_delta_1_year`: How many places a country increased or decreased with respect to last year

### Product trade at aggregated level (Year - Product Code)

If we want all products traded in 1965:
```{r yc1, eval = T}
ots_create_tidy_data(years = 1965, table = "yc")
```

If we want traded values of product "0101" (Horses, asses, mules and hinnies; live) in 1965:
```{r yc2, eval = T}
ots_create_tidy_data(years = 1965, products = "0101", table = "yc")
```

This table accepts different product codes just like `yrpc`.

Here some fields deserve an explanation:

* `pci_4_digits_product_code`: Product Complexity Index (PCI) which is detailed on [Open Trade Statistics Documentation](https://docs.tradestatistics.io/). This index is built by using just four digits product codes.
* `pci_6_digits_product_code`: Similar to the previous field but built by using just six digits product codes.
* `pci_rank_4_digits_product_code`: The rank of a product given its PCI (e.g. the highest PCI obtains the #1)
* `pci_rank_4_digits_product_code_delta_1_year`: How many places a country increased or decreased with respect to last year
